Modelling the Yelp Rating of Restaurants: Disparity of Rating Between Restaurant Cuisines and Prices

### 2024-04-19 paper/paper.qmd
Finish up paper, add map feature in

---
title: "Cuisine and Yelp Ratings"
author: 
  - Moohaeng Sohn
thanks: "Code and data are available at: [https://github.com/alexsohn1126/yelp-analysis](https://github.com/alexsohn1126/yelp-analysis)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

TODO: ADD THINGS ABOUT WHAT WE FOUND OUT AND WHY IS IT IMPORTANT

@@

We used the programming language R [@citeR], along with packages `tidyverse` [@tidyverse], `rstanarm` [@rstanarm], `jsonlite` [@jsonlite], `arrow` [@arrow], `modelsummary` [@modelsummary], `kableExtra` [@kableExtra], `here` [@here].

@@

The restaurant review dataset we will use is from Yelp. Yelp offers an academic dataset for the public to use, although it is a small subset of their massive database [@yelpdataset-faq]. The dataset is split between multiple JSON files, but we will only focus on businesses JSON. The raw businesses dataset contains 150,346 businesses. These businesses are from metropolitan areas in United States and Canada. Specifically, from metropolitan areas near Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland [@yelpdataset-faq].

@@

data %>%
  ggplot(aes(x = rating)) +
  geom_histogram(binwidth = 0.25, fill = "grey", color = "black") +
  labs(title = "Distribution of Star Ratings",
       x = "Stars",
       y = "Frequency") +
  theme_minimal()
```

@@

summary_stats <- data %>%
  summarize(
    Count = n(),
    Mean = mean(rating),
    Median = median(rating),
    SD = sd(rating),
    Minimum = min(rating),
    Maximum = max(rating)
  )

@@

# Calculate the count of restaurants per category
restaurant_summary <- data %>%
  group_by(cuisine) %>%
  summarize(restaurants_count = n()) %>%
  ungroup()

# Calculate the percentage of restaurants per category
total_restaurants <- nrow(data)
restaurant_summary <- restaurant_summary %>%
  mutate(percentage = round((restaurants_count / total_restaurants) * 100, 2))

# Arrange the result by frequency in descending order
restaurant_summary <- restaurant_summary %>%
  arrange(desc(restaurants_count))

@@

# Calculate the count of restaurants per category
restaurant_summary <- data %>%
  group_by(price) %>%
  summarize(restaurants_count = n()) %>%
  ungroup()

# Calculate the percentage of restaurants per category
total_restaurants <- nrow(data)
restaurant_summary <- restaurant_summary %>%
  mutate(percentage = round((restaurants_count / total_restaurants) * 100, 2))

# Arrange the result by frequency in descending order
restaurant_summary <- restaurant_summary %>%
  arrange(price)

@@

The exact method of how the price range was decided by Yelp is not known. We can make guesses about what kind of information they use to determine the price range, but no clear documentation is available for the price range statistics. We were able to find online discussions about the price range being decided by the check-in feature of Yelp, possibly implying that there are certain price ranges that Yelp uses to display these price ranges [@yelppricedef].

@@

We can see that nearly all cuisines has a positive effect on the rating, except for Chinese.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary(list("my model" = model))
```

```{r}
#| echo: false
#| eval: true
#| label: fig-modelresults
#| fig-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelplot(model, conf_level = 0.9) +
 # labs(x = "90% credibility interval")
```

@@

## First discussion point {#sec-first-point}

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
#| warning: false
#| message: false

data_counts <- data %>%
  group_by(cuisine, price) %>%
  summarize(count = n()) %>%
  ungroup()

# Calculate the total count of each cuisine category
cuisine_total_counts <- data_counts %>%
  group_by(cuisine) %>%
  summarize(total_count = sum(count))

# Merge the counts with the total counts
data_normalized <- data_counts %>%
  left_join(cuisine_total_counts, by = "cuisine") %>%
  mutate(normalized_count = count / total_count)

# Plot the relationship between cuisine and price range with normalized counts
data_normalized %>%
  ggplot(aes(x = cuisine, y = normalized_count, fill = factor(price))) +
  geom_bar(stat = "identity") +
  labs(title = "Relationship between Cuisine and Price Range (Normalized)",
       x = "Cuisine",
       y = "Percentage",
       fill = "Price Range") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

@@

Weaknesses and next steps should also be included.

@@

@@

plot(model, "trace") +
    theme(legend.position = "bottom")

>>>>

---
title: "Modelling the Yelp Rating of Restaurants: Disparity of Rating Between Restaurant Cuisines and Prices"
author: 
  - Moohaeng Sohn
thanks: "Code and data are available at: [https://github.com/alexsohn1126/yelp-analysis](https://github.com/alexsohn1126/yelp-analysis)"
date: today
date-format: long
abstract: "We use the Yelp dataset to explore and model how different attributes of a restaurant play a factor in the review ratings. We found there are some variables which are promising predictors for the rating of a restaurant. Cuisines such as Chinese or Mexican tended to have lower ratings, and lower priced cuisines performed worse. This study can be used to predict whether a restaurant will be successful or not."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

@@

We found out that higher the price of a restaurant, more likely it will have a higher rating, though with a couple of exceptions. And certain cuisines performed significantly worse than other cuisines. We discuss why that may be the case using graphs and tables in @sec-discussion. This study may be important to entrepreneurs whom may be interested in how these reviews may possibly be biased due to the cuisine and the price of a restaurant.

We used the programming language R [@citeR], along with packages `tidyverse` [@tidyverse], `rstanarm` [@rstanarm], `jsonlite` [@jsonlite], `arrow` [@arrow], `modelsummary` [@modelsummary], `kableExtra` [@kableExtra], `here` [@here], `leaflet` [@leaflet], `htmlwidgets` [@htmlwidgets].

@@

The restaurant review dataset we will use is from Yelp. Yelp offers an academic dataset for the public to use, although it is a small subset of their massive database [@yelpdataset-faq]. The dataset is split between multiple JSON files, but we will only focus on businesses JSON. The raw businesses dataset contains 150,346 businesses.

These businesses are from metropolitan areas in United States and Canada. Yelp claims these restaurants are from metropolitan areas near Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland [@yelpdataset-faq]. But if you map out the raw dataset as we did in @sec-dataset-map, you will see that there are no restaurants from cities such as Toronto or Montreal or Calgary. This is an error on Yelp's data documentation, but we won't be dealing with locations of restaurants directly, so we won't be heavily affected by this error.

@@

data |>
  ggplot(aes(x = rating)) +
  geom_histogram(binwidth = 0.25, fill = "grey", color = "black") +
  labs(title = "Distribution of Star Ratings",
       x = "Stars",
       y = "Frequency") +
  theme_minimal()
```

@@

summary_stats <- data |>
  summarize(
    Count = n(),
    Mean = mean(rating),
    Median = median(rating),
    SD = sd(rating),
    Minimum = min(rating),
    Maximum = max(rating)
  )

@@

# Calculate the count of restaurants per category
restaurant_summary <- data |>
  group_by(cuisine) |>
  summarize(restaurants_count = n()) |>
  ungroup()

# Calculate the percentage of restaurants per category
total_restaurants <- nrow(data)
restaurant_summary <- restaurant_summary |>
  mutate(percentage = round((restaurants_count / total_restaurants) * 100, 2))

# Arrange the result by frequency in descending order
restaurant_summary <- restaurant_summary |>
  arrange(desc(restaurants_count))

@@

# Calculate the count of restaurants per category
restaurant_summary <- data |>
  group_by(price) |>
  summarize(restaurants_count = n()) |>
  ungroup()

# Calculate the percentage of restaurants per category
total_restaurants <- nrow(data)
restaurant_summary <- restaurant_summary |>
  mutate(percentage = round((restaurants_count / total_restaurants) * 100, 2))

# Arrange the result by frequency in descending order
restaurant_summary <- restaurant_summary |>
  arrange(price)

@@

Yelp decides the price of the restaurant by crowdsourcing the price of the restaurant [@Yu_Margolin_2021]. According to @Yu_Margolin_2021, the price range of the restaurant is decided by the price brackets as shown in @tbl-prices.

```{r}
#| echo: false
#| eval: true
#| label: tbl-prices
#| tbl-cap: "Table of how Yelp Maps Price Ranges"
#| warning: false

# Make a tibble of prices, and its corresponding value in our dataset
tableData <- tibble(
  price_range_num = 1:4,
  price_brackets = c("Under $10", "$11 - $30", "$31 - $60", "Over $61")
)

# Print table
kable(tableData, align = c("c", "c"), col.names = c("Price Range Number", "Price"))
```

@@

One thing to note is that `rstanarm` takes the first factor from increasing lexicographic order and sets it as a default. So you won't see American cuisine nor price range of 1 in @tbl-modelresults and @fig-modelresults. So every coefficient listed here is how different cuisines/price ranges perform *compared to* American cuisine and price range of 1.

With that in mind, let's explore how different cuisines perform compared to American cuisine. Most cuisines are likely to increase the rating of a restaurant. However, according to our model, if a restaurant serves Chinese cuisine, it will receive a slightly lower rating than American restaurants. Also, a restaurant serving Italian or Mexican cuisines will receive roughly the same rating, as that of an American restaurant. Rest of the cuisines perform better than American cuisine.

And as for price ranges, we can see compared to restaurants with price range of 1, those with price range 2 gets significantly higher rating. The increase of rating from going from price range of 1 to 2 is larger than that of 2 to 3, and that is larger still than that of 3 to 4. In other words, we have diminishing returns as the price range increases on the rating of the restaurant.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Table of how different cuisines and price range affects the rating of a restaurant"
#| warning: false

modelsummary(list("Coefficients and Cutpoints" = model))
```

```{r}
#| echo: false
#| eval: true
#| label: fig-modelresults
#| fig-cap: "Exploring The effects of cuisines and price ranges on the rating of a Restaurant"
#| warning: false

modelplot(model, conf_level = 0.9) +
 labs(x = "Coefficient Estimates and Their 90% Credible Intervals")
```

@@

## Rating Disparity Between Cuisines {#sec-first-point}

Certain cuisines had worse effect on restaurant rating than others. Namely, Chinese, American, Italian, and Mexican cuisines had noticeably worse rating than others. While it is plausible to state that restaurants with certain cuisines are simply worse, we will try to explain this backed with research.

As for Chinese cuisine, one possible cause is the stigma around Chinese food. Chinese restaurant syndrome was a list of symptoms supposedly linked by Chinese restaurants' excessive usage of monosodium glutamate (MSG) [@Mekkodathil_Sathian_2017]. Dr. Robert Ho Man Kwok first coined the term "Chinese Restaurant Syndrome" back in April of 1968, sending a letter to the New England Journal of Medicine [@LeMesurier_2017]. This stigma against Chinese restaurants could still be remaining today, leading to a lower rating.

American, Italian, and Mexican cuisine's bad performance could be explained by the demographic of the restaurants. All the restaurants in the dataset are in United States and Canada. All 3 aforementioned cuisines have very close ties with these countries. American cuisine is from United States. Italian cuisine is from Italy, but a lot of Italians are living in North America. Around 16 million Italian descendants are living in United States [@USCensus], and 1.5 million in Canada [@canadaCensus]. Mexico shares a border with United states, therefore a ton of cultural influences were made in southern United States by Mexicans. What we propose is that people rate restaurants that serve foods from their own culture harshly. For example, people who have Mexican parents or has been raised with Mexican culture will have higher standards for Mexican food.

We can try attributing this to the fact that different cuisines will have different price ranges. Therefore one can argue that those cuisines with higher proportion of low price restaurants will be more likely to have lower ratings. @Fig-cuisine-price-rel shows each cuisine's price range proportions. It shows that while Chinese and Mexican restaurants have high proportion of restaurants with price range of 1, but it is not the case for American and Italian cuisine.

```{r}
#| label: fig-cuisine-price-rel
#| fig-cap: "Relationship between Cuisine and Price Range (Normalized)"
#| echo: false
#| warning: false
#| message: false

data_counts <- data |>
  group_by(cuisine, price) |>
  summarize(count = n()) |>
  ungroup()

# Calculate the total count of each cuisine category
cuisine_total_counts <- data_counts |>
  group_by(cuisine) |>
  summarize(total_count = sum(count))

# Merge the counts with the total counts
data_normalized <- data_counts |>
  left_join(cuisine_total_counts, by = "cuisine") |>
  mutate(normalized_count = count / total_count) |>
  arrange(count)

# Plot the relationship between cuisine and price range with normalized counts
data_normalized |>
  ggplot(aes(x = cuisine, y = normalized_count, fill = factor(price))) +
  geom_bar(stat = "identity") +
  labs(x = "Cuisine",
       y = "Percentage",
       fill = "Price Range") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Another interesting insight is that these 4 cuisines we have listed as "worse" than others are the top 4 cuisines excluding "Other" category in our dataset (Check @tbl-category-summary). Perhaps more restaurants there are for certain cuisines, people's expectations are higher, leading to lower rating for "average" experiences in those restaurants. @Fig-cuisine-ratings shows the relationship between the number of restaurant and the rating that a cuisine has. We can clearly see there is a downward trend as number of restaurants increase. This could have occurred because people associate rare cuisine as being better. Also, there is a possibility that the owners that would care about their Yelp categories are also more likely to care about the service and the food of the restaurant.

```{r}
#| label: fig-cuisine-ratings
#| fig-cap: "Average Rating vs. Number of Restaurants by Cuisine"
#| echo: false
#| warning: false
#| message: false

# Get average rating per cuisine
cuisine_summary <- data %>%
  group_by(cuisine) %>%
  summarize(num_restaurants = n(), avg_rating = mean(rating))

# Plot number of restaurants vs average rating
ggplot(cuisine_summary, aes(x = num_restaurants, y = avg_rating)) +
  geom_point() +
  geom_text(aes(label=cuisine), nudge_y = 0.02, size = 3, angle = 25) +
  labs(x = "Number of Restaurants", y = "Average Rating") +
  theme_minimal()
```

## Price Range and Rating

We also observed that higher price range of a restaurant means the restaurant will higher ratings. We have expected this in the model section, but the diminishing returns of ratings as price increases was not expected. We can plot out the relationship between the price and the rating by cuisine to further illustrate this idea. @Fig-rating-price shows that the lower rating cuisines do have lower average prices. But there are cuisines, such as Caribbean or Vietnamese that has a high rating, yet with lower average prices. The figure shows that for some cuisines, their price range doesn't affect their ratings.

```{r}
#| echo: false
#| eval: true
#| label: fig-rating-price
#| fig-cap: "Average Rating vs. Average Price by Cuisine"
#| warning: false

cuisine_summary <- data %>%
  group_by(cuisine) %>%
  summarize(avg_price = mean(as.numeric(factor(price), ordered = TRUE)),
            avg_rating = mean(rating))

# Plot
ggplot(cuisine_summary, aes(x = avg_price, y = avg_rating, label = cuisine)) +
  geom_point() +
  geom_text(aes(label = cuisine), nudge_x = 0.05, nudge_y = 0.05, size = 3, angle = 45, hjust = 1) +  # Add text labels with a slight nudge and angle
  labs(x = "Average Price", y = "Average Rating",) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This could have possibly been caused by simply those with lower prices cannot afford decent service and food, but that still does not explain the low average rating of Italian and American restaurants which have higher costs yet near the same rating as Mexican and Chinese restaurants. Another way this could have been caused is because, again, people rate rarer cuisines higher than common ones. So perhaps in later studies, more research can be done with those rarity in mind.

@@

A weakness that we have observed is that there may not be enough variables in our model to correctly model the rating. We could have added more factors, such as how rare the cuisine is in the dataset, or the number of reviews. We might have enough information to perform post-stratification with the number of ratings for a restaurant, but that would require a reliable, bigger dataset on reviews of restaurants which we do not have access to.

Another limitation could have been in how we clean our data. We selected businesses that had restaurant-related categories in Yelp, but there is likely to be many businesses which forgot to add the restaurant related categories on Yelp, or simply does not care. Therefore, filtering restaurants by categories is unreliable.

The data itself is also a limitation. The data only contains restaurants from big cities, such as Toronto, Vancouver, and Las Vegas. This could have caused a skew in our dataset, over-representing restaurants in bigger cities, and not those in the smaller, suburban communities. Not only that, the ratings of restaurants were given in 0.5 increments which meant it isn't the most accurate rating we could have gotten.

Next steps to take are refining this model to take account of different variables that were missed in this paper and improving the predictability of the model. Perhaps with a newer database, this study shall be repeated too.

@@

# Cleaned Dataset Interactive Map {#sec-dataset-map}

@@

plot(model, "trace") +
    theme_minimal() +
    theme(legend.position = "bottom")
